---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---


## Loading and preprocessing the data
```{r, echo=TRUE}
dat.ori <- read.csv(file="activity.csv", header = TRUE ,na.strings="NA" )
str(dat.ori)
head(dat.ori)

dat.ori$date <-  as.Date(dat.ori$date)
head(dat.ori)
```


## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library(dplyr)
dat.day <- dat.ori %>%
        select(date, steps) %>%
        group_by(date) %>%
        summarise_each(funs(sum))
head(dat.day)

hist(dat.day$steps,
     main="Total Number of steps taken each day",
     xlab="Total number of steps")
dev.copy(png,  file  =  "figure/plot1.png")
dev.off()
```

```{r, echo=TRUE}
summary(dat.day$steps)
```

The mean total number of steps taken per day is 10770 steps. 
The median total number of steps taken per day is 10760 steps. 

## What is the average daily activity pattern?
# Time series plot of the average number of steps taken
```{r, echo=TRUE}
dat.int <- dat.ori %>% 
        select(steps, interval) %>% 
        group_by(interval) %>%
        summarise_each(funs(mean(., na.rm = TRUE)))

plot(dat.int$interval,
     dat.int$steps,
     type="l",
     main =" average number of steps taken  across all days in each interval",
     xlab ="interver identifier",
     ylab="average steps across all days")
dev.copy(png,  file  =  "figure/plot2.png")
dev.off()
```
# The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}
max.avgsteps <- max(dat.int$steps, na.rm = TRUE)
subset(dat.int, steps == max.avgsteps)
```
The 835 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

## Imputing missing values
```{r, echo=TRUE}
dat.miss <- dat.ori[!complete.cases(dat.ori),]
nrow(dat.miss)
```
The total number of missing values in the dataset is 2304

```{r, echo=TRUE}
dat.imput <- merge(dat.miss,dat.int, by="interval")
head(dat.imput)
dat.imput <- dat.imput[, c ("interval", "date", "steps.y")]
names(dat.imput) <- c("interval", "date", "steps")
head(dat.imput)
nrow(dat.imput)

nrow(dat.ori)
nrow(dat.ori) - nrow(dat.imput)

dat.complet<- na.omit(dat.ori) 
nrow(dat.complet)
dat.imput <- rbind(dat.imput, dat.complet)
nrow(dat.imput)
```

# 7 Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo=TRUE}
dat.day2 <- dat.imput %>%
        select(date, steps) %>%
        group_by(date) %>%
        summarise_each(funs(sum))
hist(dat.day2$steps,
     main="Total Number of steps taken each day ( after missing data imputation) ",
     xlab="Total number of steps")
dev.copy(png,  file  =  "figure/plot3.png")
dev.off()
summary(dat.day$steps)
summary
```
After missing data imputation, he mean and median total number of steps taken per day are  10760  and  10770 , 
They do not differ from the estimate from the first part of the assignment. 
Since missing data was replaced by the mean of that interval.
I think missing data imputation will have an influence of the standard error.
Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
weekday1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
dat.imput$wday <- factor((weekdays(dat.imput$date) %in% weekday1),
                    levels =c(FALSE, TRUE), labels=c("weekend", "weekday") )
levels(dat.imput$wday)

dat.int2 <- dat.imput%>%
        select(steps,interval, wday) %>%
        group_by(wday, interval) %>%
        summarise_each(funs(mean))

library(ggplot2)
ggplot(dat.int2, aes(x=interval, y=steps)) +
        geom_line(colour="blue") +
        facet_grid(wday ~. )
dev.copy(png,  file  =  "figure/plot4.png")
dev.off()
```

